### 2016-02-06 scrap/costs.Rmd
more integration into remake

---
title: "Costs"
author: "John"
date: "1 February 2016"
output: html_document
---

Costs of fighting

I was thinking something like

$$ E_{costs} = \frac {E_{loser max}} {1 + exp(-(\frac{m_{winner} - m_{loser}}{k} - m_0)}$$

Where the energy cost of fighting is deducted from both sides.
$E_0$ describes the center, ie the cost to two evently matched opponents with $E_0 = 0$ is 50%

```{r}
library(rasterVis)
loser_masses <- seq(1, 100, by = 0.1)
winner_masses <- seq(1, 100, by = 0.1)
e_0 <- -0.5
E_loser_max <- 100

costs <- function( m_loser, m_winner, k){
    E_loser_max / (1 + exp(-((m_winner - m_loser) / k - e_0)))
}
```
costs_scaling <- 25:
```{R}
cost_scaling <- 100 # k 
r <- raster(outer(loser_masses, winner_masses, function(a,b){costs(a, b, cost_scaling)}), 
              xmn = 0,
              xmx = 100,
              ymn = 0,
              ymx = 100)

plot(r, 
  main = "Cost of Fighting k = 25", 
  xlab = "loser's mass",
  ylab = "winner's mass")
```

costs_scaling <- 50:
```{R}
costs_scaling <- 1
r <- raster(outer(loser_masses, winner_masses, function(a, b){costs(a, b, cost_scaling)}),
              xmn = 0,
              xmx = 100,
              ymn = 0,
              ymx = 100)

plot(r,
  main = "Cost of Fighting k = 1", 
  xlab = "loser's mass",
  ylab = "winner's mass")

```