### 2016-02-07 ms.tex
comparison between energy budgets and cost functions

\documentclass[a4paper,11pt]{article}
\usepackage[osf]{mathpazo}
\usepackage{ms}
\usepackage[]{natbib}
\raggedright

\newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
\usepackage{graphicx}

\title{Toward dynamic models of combat or lack thereof in animal mating}
\author{
* John Wilshire$^1$, Will Cornwell$^1$ , Daniel Falster$^2$, Michael Kasumovic$^1$, \\
Daniel Noble$^1$,$^1$ Loic Thibaut$^1$}
\affiliation{
*final list and order undecided\\
$^1$ University of NSW\\
$^2$ Macquarie University\\
}
\date{}

\bibliographystyle{mee}

\usepackage[title,titletoc,toc]{appendix}

\mstype{Research Article}
\runninghead{A new framework for fighting}
\keywords{}

\begin{document}
\mstitlepage
\noindent
% \doublespacing
% \linenumbers

\section{Summary}
The diversity of animal mating systems is astounding. In some of these
systems, very costly combat behaviour -- among males, among females, or
both -- is a feature of the mating process.  In other systems, resources
are divided among individuals in an entirely pacific process.  Can we
understand why? Animal combat strategies likely emerge from trade-offs
in investment in growth, mate seeking, and information gathering.
Willingness to engage in combat is a trait that evolves based on the
fitness landscape, which itself changes depending on both the
environment and the strategies of other individuals.  Using recently
developed methods for modelling dynamic fitness landscapes, we examine:
(1) why combat behaviours arise, (2) under what conditions combat
behaviours are evolutionarily stable, and (3) when different combat
strategies co-exist.  We hypothesize that the reliability and
"public-ness" of information is an important feature driving combat or
lack thereof in many animal systems.

\section{Introduction}

Evolution of animal personalities: \citep{Wolf-2007,Wolf-2012} show can have
coexistence of risky, explorative strategies and risk-averse strategies.

Animal personalities linked to other life history traits: \citep{Biro-2008}

Individual-based models of natural selection: \citep{MGonigle-2012}

\section{Methods}

% Verbal description of model

We consider a population of males competing for mates. The population is one with non-overlapping generations, and within each generation there is a predefined annual cycle. In spring, up to $K$ individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies. The short duration of mating period is such that each male and female only mates once. The population then re-establishes from eggs the following spring.

The reproductive success of males in the population is determined via their ability to compete for and hold $N$ nests. Nests differ in quality (potential number of offspring per nest per year). The male population is divided into two groups: immature and mature males. Immature males devote all their efforts towards feeding, increasing their size. Males mature with a probability defined by their size and a maturation trait. Once they enter the mature population, males stop feeding and focus solely on obtaining nests. Nests can be vacant or occupied and males are either ``searching'' or ``occupying'' a nest. Males differ in the rate at which they search, and thus the rate at which they encounter nests. Empty nests can be immediately occupied with no additional cost. Filled nests can only be obtained by entering into and winning a contest. Once occupied, males can choose to occupy a nest or continue searching. At the end of each season, each male has a fitness defined by the reproductive value of the nest it holds, $R_i$.

\subsection{Habitat}

Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$. The distribution of nest qualities is then set according to the distribution
\begin{equation} \label{eq:pdf_R}
    Pr(R_i = x) =\bar{R} \, p(R).
\end{equation}
The total number of male offspring produced in each generation is then
\begin{equation} \label{eq:pdf_R}
    K = \bar{R} N.
\end{equation}

\subsection{Immature phase}
% TODO update this to reflect what happens in the model
% ie pulling the maturation times from the logistic cdf with location and parameters given

Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
\frac{{\rm d} m}{{\rm d} t} = a \, m^ b
\end{equation}
where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
\begin{equation} \label{eq:growth}
m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}},
\end{equation}

while the time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
\begin{equation} \label{eq:tau}
\tau(m_0, m) = \frac1{a(1-b)}\left(m ^{1-b} - m_0 ^{1-b}\right).
\end{equation}

If we assume a constant mortality rate $k_i$ during the immature phase, the probability of males surviving to time $t$ is given by
\begin{equation} \label{eq:surv_immature}
S(t) = \exp(-k_i \, t).
\end{equation}

Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
\begin{equation}\label{eq:allocation}
Pr({\rm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\rm mat}\right)\right)},
\end{equation}
where $m_{\rm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.

\subsection{Mature phase}
When a male matures it is given an energy budget linearly proportional to its mass, this is a parameter into the model.\\

\subsection{Searching Males}
When males mature they begin searching for a females nest.

The area that a male $m$ with speed trait $v$ and radius trait $r$ can search in one time step is
\begin{equation}
    2 \cdot r \cdot v
\end{equation}
See Fig. \ref{fig:area searched}.

We let the total area enclosing all males and all nests be a parameter, $A_{total}$. We assume spatially random nests.

The encounter rate $\varepsilon$ for an individual is:
\begin{equation} \label{eq:encounter rate}
    \varepsilon = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
\end{equation}
\citep{Gurarie2012}

We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females mature.
When an event occurs we select a nest randomly and then the male will either occupy it or contest another male for it.

When a male is kicked out of a nest we discard the old events and generate new ones.

The energy used by a male with speed $v$ and radius $r$ is 
\begin{equation}
    E(r,v) = k_{search} \cdot r \cdot v
\end{equation}
where $k_{search}$ is a constant.

\clearpage
\subsection{Contest Algorithm}
See Fig. \ref{fig:contests}.
For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
The first stage is the display stage, the second phase is the fight stage.
each individual $i$ has mass $m$. At the time of the contest it also has an amount of energy $E_{current}$.  
The aggression of the individual is a function of $E_{current}$:

\begin{equation}\label{eq:aggression}
a_i = f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
\end{equation}

$L$ is a scaling parameter.

Contests can be escalated to different degrees with increasing energy costs for both participants. 
Either participant may abandon the contest at any moment, they do this by choosing not to escalate.
We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantaneously.

Traits $K$ and $E_{o}$ are heritable and potentially under selection.

The probability of the attacker escalating to  is:

\begin{equation}\label{eq:attacker_escalating}
Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
\end{equation}

Similarly the probability of the defender escalating is:

\begin{equation}\label{eq:defender_escalating}
          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
\end{equation}

The energetic cost of escalating is:
\begin{equation}\label{eq:defender_escalating}
E_{spent}=equation you guys figured out with all the \lambda 's
\end{equation}

The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.

If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
After the contest energy costs are deducted from both the winner and the loser.
If a death occurs ($E \leq 0$) then x is removed from the individual pool.

\clearpage

\subsection{Evolutionary dynamics}

A new individual will inherit the fathers genes,
There is a global mutation rate parameter which is the chance that a trait will not be inherited perfectly but instead will be mutated.
This means adding some normally distributed noise with $\mu = 0$ and $\sigma$ varying from trait to trait.

\begin{enumerate}
    \item Abandon Rate, influences RR abandonment rate (not yet implemented)
    Trade offs:
    \begin{itemize}
        \item more likely to abandon a nest and not discover another before the females mature
    \end{itemize}

    \item Aggressiveness, determined by two inheritable traits $k$, $E_0$
    Trade offs:
    \begin{itemize}
        \item more likely to engage in fights with a low probability of winning
    \end{itemize}

    \item search radius, the radius of a searching male (can be thought of as line of sight), See Fig. \ref{fig:area searched}.
    Trade offs:
    \begin{itemize}
        \item higher metabolic cost of search
    \end{itemize}

    \item search speed, the distance that a searching male will cover in one time step, see Fig. \ref{fig:area searched}.
    Trade offs:
        \begin{itemize}
            \item higher metabolic cost of search
            \item higher predation rate
        \end{itemize}
    \item maturation time, the time at which to mature is inherited as well but with a constant level of noise
\end{enumerate}

\clearpage

\section{Results} % (fold)
\label{sec:results}

\begin{figure}[h!]
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
    \caption{Flat costs. Percent energy expenditure of cohort at generation 499}
    \label{fig:energy pc}

    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot.pdf}
    \caption{ Flat costs Total energy expenditure of cohort at generation 499}
    \label{fig:energy total}
\end{figure}
As patch area increases the number of fight importunities decreases,
This can probably explain the strong decrease in energy spent fighting.but I have no idea, early days yet.
\clearpage

\begin{figure}[h!]
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_var.pdf}
    \caption{Variable costs. Percent energy expenditure of cohort at generation 499}
    \label{fig:var energy pc}
    \centering
    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot_var.pdf}
    \caption{Variable costs, Total energy expenditure of cohort at generation 499}
    \label{fig:var energy total}
\end{figure}

The results for variable costs are different, why? from here on all graphs are from the flat cost data...

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_density.pdf}
    \caption{a density plot of the aggression parameter k at max energy}
    \label{fig:k density}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_density.pdf}
    \caption{a density plot of the aggression parameter $e_0$ at max energy}
    \label{fig:e_0 density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_history.pdf}
    \caption{a line(ish) plot of the aggression parameter $k$ at max energy}
    \label{fig:k history}

    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_history.pdf}
    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy}
    \label{fig:e_0 history}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=8cm,keepaspectratio]{figures/aggression_density.pdf}
    \caption{a density plot of aggression at max energy. I am trying to get it to be multiple plots}
    \label{fig:aggression density}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/num_winners.pdf}
    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
    \label{fig:num winners}
\end{figure}

\clearpage
% section results (end)results

I am not sure if this is the best way to do this, the tables in the plant paper are more aesthetic.\\

\begin{table}[h!]
    \caption{Variable names and definitions in the model.}
    \centering
    \begin{tabular}{c | c | l }
        \hline
        Symbol & unit & Description\\
        \hline
        \hline
        $K$ & & number of individual males \\
        $N$ & & number of nests \\
        $R_i$ & & reproductive value of nest of nest $i$\\
        $\bar{R}$ & & average reproductive value of nest\\
        \hline
        $t_m$ & & the time at which the females begin to mature\\
        $S(t)$ = exp($-k_it$) & & The probability of an individual surviving to time t.\\
        \hline
        $m$ & $g$ & mass\\
        $m_mat$ & $g$ & The size at which 50\% of the population is mature.\\
        $E$ & kJ & kiloJoules\\
        \hline
    \end{tabular}
\end{table}

\section{Figures}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/area_searched}
\caption{the area searched by a male in interval $dt$}
\label{fig:area searched figure}
\end{figure}
\clearpage

\bibliography{refs}

\end{document}