### 2016-02-18 scrap/evolution.Rmd
new plots and evolution rmd

---
title: "Evolution plots"
author: "John Wilshire"
date: "18 February 2016"
output:
  pdf_document:
    fig_crop: no
    fig_height: 4
  html_document: default
---

```{r, echo=FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("../R/load_datasets.R")
source("../R/make_plots.R")
```

These are roughly equivilant to the density plots but I feel the density plot is a little bit too messy...

```{R}
expf10_pop <- load_population("../data/exp_f10/")
expf15_pop <- load_population("../data/exp_f15/")
linf10_pop <- load_population("../data/lin_f10/")
population_plots <- function (what){
  print(pop_values(expf10_pop, what, paste("expf10_pop", what)))
  print(pop_values(expf15_pop, what, paste("expf15_pop", what)))
  print(pop_values(linf10_pop, what, paste("linf10_pop", what)))
}

population_plots("e_0")
population_plots("k")
population_plots("2 * radius * speed")

```

These plots trace how parameters evolve through generations with coloured by patch area...

```{r}
expf10 <- load_trait_history("../data/exp_f10/")
expf15 <- load_trait_history("../data/exp_f15/")
linf10 <- load_trait_history("../data/lin_f10/")

evolution_plots <- function (what){
  print(trait_evolution_plot(expf10, what, paste("expf10", what)))
  print(trait_evolution_plot(expf15, what, paste("expf15", what)))
  print(trait_evolution_plot(linf10, what, paste("linf10", what)))
}

evolution_plots("mean_k")
evolution_plots("std_k")
evolution_plots("std_e_0")
evolution_plots("num_winners")
evolution_plots("mean_maturation")
evolution_plots("contest_energy")
evolution_plots("search_energy")
evolution_plots("occupying_energy")
evolution_plots("mean_exploration")
evolution_plots("std_exploration")
```