### 2016-03-15 scrap/clustering.rmd

---
title: "Clustering"
author: "John Wilshire"
date: "15 March 2016"
output: pdf_document
---

```{r, echo=FALSE}
library(fpc)

library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")
```

the pamk function finds the ideal k for k means.

```{R}

exp_100 <- load_population("~/Documents/contests_model/data/cm_100/")

exp_1000 <- exp_100[exp_100$patch_area %in% seq(1000,30000, by = 500),]

pamk_clusters <- ddply(exp_1000, .variables = .(patch_area), 
          .fun = function (df){
            m <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
            cluster <- m$pamobject$clustering
            mediod <- m$pamobject$medoids[cluster]
            distance <- m$pamobject$clusinfo[cluster, "diameter"]
            nc <- max(cluster)
            unique(data.frame(cluster, mediod, distance, nc))
          })

ggplot(pamk_clusters, aes(x = patch_area, y = mediod)) +
  geom_errorbar(aes( ymin = mediod - (distance/2), ymax = mediod + (distance/2))) +
  geom_point(aes(col=as.factor(nc))) + 
  ylab("log(beta)") + 
  ggtitle("Clustered using pamk") + 
  scale_colour_discrete(name="Number of Clusters")

qplot(data = exp_1000, maturation_time, 
      fill=as.factor(patch_area), 
      main ="maturation times of winners")

qplot(data = exp_1000, mass, fill=as.factor(patch_area), main ="mass distrobution of winners, mass center of inital population is 4")

qplot(data = exp_1000, x = patch_area, y = mass)

```