### 2016-03-18 scrap/contest_data.Rmd

---
title: "actual contest data"
author: "John Wilshire"
date: "18 March 2016"
output: html_document
---

```{R, echo = FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

contests <- data.frame(fromJSON("~/Documents/contests_model/data/03_18_05_08_07__c7a73359-a904-4c5c-ba5b-189dff605a90.json")[c("contests", "parameters")])
colnames(contests) <- sub("^[^.]*.","", colnames(contests))
```

Seems pretty reasonable...
```{R} 
ggplot(contests, aes(x = def_mass, y = atk_mass)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("mass of the attacker") + 
  xlab("mass of the defender") + 
  scale_colour_discrete(name="Defender Wins")
```

```{R}
ggplot(contests, aes(y = atk_start_energy , x = def_start_energy )) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("initial energy of the attacker") + 
  xlab("initial energy of the defender") + 
  scale_colour_discrete(name="Defender Wins")

```

It looks like a lot of the contests end in death
```{R}
ggplot(contests, aes(y = atk_end_energy , x = def_end_energy )) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("final energy of the attacker") + 
  xlab("final energy of the defender") + 
  scale_colour_discrete(name="Defender Wins")

```

The intensity function is $c^2$
```{R}
ggplot(contests, aes(x = def_commit, y = atk_commit)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("commitment of the attacker") + 
  xlab("commitment of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_beta, y = atk_beta)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("beta trait of the attacker") + 
  xlab("beta trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_beta, y = atk_beta)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("beta trait of the attacker") + 
  xlab("beta trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_alpha, y = atk_alpha)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("alpha trait of the attacker") + 
  xlab("alpha trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

qplot(contests$fight_cost, fill= contests$defence_winner)

```