### 2016-03-23 scrap/clustering.rmd
r and v can be fixed

---
title: "Clustering"
author: "John Wilshire"
date: "15 March 2016"
output: pdf_document
---

```{r, echo=FALSE}
library(fpc)

library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")
```

the pamk function finds the ideal k for k means.

```{R}

exp_100 <- load_population("~/Documents/contests_model/data/cm_100/")

cm_6k <- load_population("~/Documents/contests_model/data/cm_100/")
clustered_plot(cm_6k, what = "e_0", ylab = "beta")

exp_1000 <- exp_100[exp_100$patch_area %in% seq(1000,30000, by = 500),]

pamk_clusters <- ddply(exp_100, .variables = .(patch_area), 
          .fun = function (df){
            c1 <- pamk(df$k, krange = 1:3, alpha = 1e-12)
            k_cluster <- c1$pamobject$clustering
            k_mediod <- c1$pamobject$medoids[k_cluster]
            k_distance <- c1$pamobject$clusinfo[k_cluster, "diameter"]
            k_nc <- max(k_cluster)

            c2 <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
            e_0_cluster <- c2$pamobject$clustering
            e_0_mediod <- c2$pamobject$medoids[e_0_cluster]
            e_0_distance <- c2$pamobject$clusinfo[e_0_cluster, "diameter"]
            e_0_nc <- max(e_0_cluster)

            data.frame(k_cluster, k_mediod, k_distance, k_nc, e_0_cluster, e_0_mediod, e_0_distance, e_0_nc)
          })

ggplot(data = pamk_clusters, aes(x = patch_area)) + 
  geom_point(aes(y = k_nc, col = as.factor(e_0_nc)))

pamk_1000 <- pamk_clusters[pamk_clusters$patch_area == 1000, ]

ggplot(data = exp_100[exp_100$patch_area == 1000, ], aes(x = k)) + 
  geom_point(aes(y = e_0))
  geom_polygon

ggplot(pamk_clusters, aes(x = patch_area, y = mediod)) +
  geom_errorbar(aes( ymin = mediod - (distance/2), ymax = mediod + (distance/2))) +
  geom_point(aes(col=as.factor(nc))) + 
  ylab("beta") + 
  ggtitle("Clustered using pamk") + 
  scale_colour_discrete(name="Number of Clusters")

qplot(data = exp_1000, maturation_time, 
      fill=as.factor(patch_area), 
      main ="maturation times of winners")

pamk(exp_100[exp_100$patch_area == 4600,]$k, krange = 1:3, alpha = 1e-12)

```