### 2016-03-23 scrap/contests_accross_patch_areas.rmd

---
title: "contests accross patch areas"
author: "John Wilshire"
date: "21 March 2016"
output: pdf_document
---

contests_accross_patch_areas.rmd

6k generations, mutation sd 0.05, rate 0.1

```{R}
library(jsonlite)
library(plyr)
library(fpc)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

```
```{R}
all_contests <- load_contests("~/Documents/scrap/contest_data/contest_logging/")
winners <- load_population("~/Documents/scrap/contest_data/contest_logging/")

contests_summary <- ddply(all_contests, .(patch_area), summarise,
                          number_of_fights = length(fight_cost),
                          mean_fight_cost = mean(fight_cost),
                          def_win_rate = mean(defence_winner),
                          atk_death_rate = mean( atk_end_energy  <= 0),
                          def_death_rate = mean( def_end_energy  <= 0))

qplot(data = contests_summary, x = patch_area, y = mean_fight_cost, main = "mean fight cost")
qplot(data = contests_summary, x = patch_area, y = def_win_rate, main = "defender win rate")
qplot(data = contests_summary, x = patch_area, y = atk_death_rate, main = "atacker death rate")
qplot(data = contests_summary, x = patch_area, y = def_death_rate, main = "defender death rate")

```
```{R}
# lots of deaths?
qplot(data = contests_summary, x = patch_area, y = def_death_rate + atk_death_rate)
contests_summary$density <- 500 / conteasts_summary$patch_area
ggplot(contests_summary, aes(x = density, y = number_of_fights)) + geom_point() + scale_y_log10() + scale_x_log10() + geom_smooth(method = lm)

summary(lm(contests_summary$number_of_fights~log10(contests_summary$density)))

ggplot(all_contests[all_contests$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = fight_cost)) +
  geom_violin()  
# the bars are the max - min in that cluster ymin = mediod - (max - min)/2
# this allows for negative values where there arent acually any negative values if the max is an outlier

clustered_plot(dff = all_contests, what = "fight_cost", ylab = "fight_cost")

summary(all_contests$fight_cost)
```

```{R}
# one generation with low patch area / high density

low_patch_area <- all_contests[all_contests$patch_area == 2200,]
high_patch_area <- all_contests[all_contests$patch_area == 25000,]

qplot(data = low_patch_area, x = def_mass, y = atk_mass, col = defence_winner, main = "low patch area")

qplot(data = high_patch_area, x = def_mass, y = atk_mass, col = defence_winner, main = "low patch area")

```

Here is some other plots 
```{R}
clustered_plot(dff = winners, what = "e_0", ylab = "beta")

clustered_plot(dff = winners, what = "k", ylab = "alpha")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = e_0)) +
  geom_violin() + 
  ggtitle("survivors beta") + 
  ylab("beta")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 500),],
       aes(x = factor(patch_area), y = k)) +
  geom_violin() + 
  ggtitle("survivors alpha") + 
  ylab("alpha")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 500),],
       aes(x = factor(patch_area), y = mass)) +
  geom_violin() + 
  ggtitle("survivors mass") + 
  ylab("mass")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = 2 * speed * radius)) +
  geom_violin() + 
  ggtitle("exploration trait") + 
  ylab("2 * speed * radius")

```