### 2016-03-23 scrap/last_gen_energy.rmd

---
title: "energy_at_max_gen"
author: "John Wilshire"
date: "18 March 2016"
output: html_document
---

```{r}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

ts_6k <- load_trait_history("~/Documents/contests_model/data/6k/")

last_gens <- ddply(ts_6k, .variables = .(patch_area),
                   function(df) {
                     max_gen <- max(df$generation)
                   })

```