### 2016-04-04 scrap/generation_comparison.Rmd

---
title: "A Comparison of Different Generation lengths"
author: "John"
date: "19 February 2016"
output: html_document
---
I ran one simulation stepping by 100 for 500 generations.
and another stepping by 250 for 2000 generations.

```{r}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")
cm_100 <- load_population("~/Documents/contests_model/data/cm_100/")
cm_2k <- load_population("~/Documents/scrap/contest_data/6k/")
low_logit <- load_population("~/Documents/contests_model/data/low_logit/")
low_logit_100 <- load_population("~/Documents/contests_model/data/low_logit_100/")
repeated <- load_population("~/Documents/scrap/contest_data/repeated/")

# get upper and lower bounds of trait values
get_limits <- function (what, dfs){
  return(c(min(dfs[what]), max(dfs[what])))
}

population_plots <- function (x, y, limits){
  print(
    pop_values(cm_100, x = x, y = y, paste("cm_100",x, y)) + 
      scale_y_continuous(limits = limits)
  )
  print(
    pop_values(cm_2k, x = x, y = y, paste("cm_6k",x, y)) + 
      scale_y_continuous(limits = limits)
  )
    print(
    pop_values(low_logit_100, x = x, y = y, paste("low_logit_100",x, y)) + 
      scale_y_continuous(limits = limits)
  )
      print(
    pop_values(low_logit, x = x, y = y, paste("low_logit",x, y)) + 
      scale_y_continuous(limits = limits)
  )
}
e_0_limits <- get_limits("e_0", rbind(cm_2k, cm_100, low_logit_100, low_logit))
k_limits <- get_limits("k", rbind(cm_2k, cm_100, low_logit_100, low_logit))
```

The low_logit graphs are those that have a the coef in the logistic of swapping = 1 compared to 10 in the cm ones,

some interesting co-existance,
but the same scattering problem...

```{r, dev='png'}
population_plots("log(N / patch_area)", "e_0", e_0_limits)
population_plots("log(N / patch_area)", "k", k_limits)

population_plots("patch_area", "e_0", e_0_limits)
population_plots("patch_area", "k", k_limits)
```