### 2016-04-04 scrap/stabilisation.Rmd

---
title: "Stabilisation of Traits"
author: "John Wilshire"
date: "29 February 2016"
output: pdf_document
fig.show: hold
---

```{r, tidy=FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

#traits_cm_100 <- load_trait_history("~/Documents/contests_model/data/cm_100/")
traits_cm_2k <- load_trait_history("~/Documents/contests_model/data/cm_2k/")

stable_what_where <- function (df, what, where){
  print(
    trait_evolution_plot(df[df$patch_area %in% where,], what, "title")

  )
}

stable_what_where(traits_cm_2k, "std_e_0", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "mean_e_0", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "std_k", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "mean_k", seq(0, 30000, by = 5000))
```